---
title: 'Multiple Comparisons: Homework - 2'
output:
  pdf_document: default
  github_document: 
    pandoc_args: --webtex
  html_notebook: default
---


Submitted by: 
=============

Eyal Bar-Natan - 207630658

Omer Shubi - 312236219


# Question 1


## Part A


We perform $iterations=5,000$ simulations, in which each time we randomly pick $m=5$ values out a Uniform Distribution ($U\sim Uni[0,1]$).
In each simulation, out of the picked values we select the one with the minimum value. The result of these simulations is stored in $vecmin\_m$.

 
```{r}
iterations <- 5000
```

```{r}
m <- 5
mat_5 <- replicate(iterations, runif(m,0, 1))
vecmin_5 <- apply(mat_5, 2, min)
```

and similarly with $m=20$:

```{r}
m <- 20
mat_20 <- replicate(iterations, runif(m,0, 1))
vecmin_20 <- apply(mat_20, 2, min)
```

and $m=100$:

```{r}
m <- 100
mat_100 <- replicate(iterations, runif(m,0, 1))
vecmin_100 <- apply(mat_100, 2, min)
```

### 1.

The proportion of $U_{(1)}<0.05$ when $m=5$:

```{r}
length(vecmin_5[vecmin_5<0.05])/iterations
```

Similarly to section $\aleph$ in H.W. 1 - in which $p_{value}\sim Uni[0,1]$, and we were requested to retrieve the minimal $p_{value}$.

This is exactly what we are asked to find in this question as well, 
by defining $U_i={p_{value}}_i\sim Uni[0,1]$ and $U_{(1)} = min(U_1,U_2,\ldots,U_m)$
we are dealing with the same thing.

### 2.

Histogram of $U_{(1)}$ with $m=5$:
```{r}
hist(vecmin_5, xlim=c(0,1))
```
Histogram of $U_{i}$ with $m=5$:

```{r}
hist(mat_5, xlim=c(0,1))
```


By running the ecdf function we can visual the cdf of both functions clearly. 

```{r}
plot(ecdf(vecmin_5),col='red', main='CDF of Uni[0,1] and U(1)', xlim=c(0,1))
lines(ecdf(mat_5),col='green')
legend(0, 1, legend=c(expression(U[(1)]), 'Uni[0,1]'), col=c("red", "green"), lty=1, cex=0.8)
```
As can be seen from both the histograms and the cdf functions,

it is clear that $U_{(1)}$ is stochastically smaller ($\prec$) than $Uni[0,1]$.

### 3.


Histogram of $U_{(1)}$ with $m=5$:

```{r}
hist(vecmin_5, xlim=c(0,1))
```
Histogram of $U_{(1)}$ with $m=20$:

```{r}
hist(vecmin_20, breaks = 20, xlim=c(0,1))
```
Histogram of $U_{(1)}$ with $m=100$:

```{r}
hist(vecmin_100, xlim=c(0,1))
```

```{r}
plot(ecdf(vecmin_5),col='red', main='CDF of U(1) with different m values', xlim=c(0,0.65))
lines(ecdf(vecmin_20),col='green')
lines(ecdf(vecmin_100),col='blue')

legend(0.5, 0.5, legend=c(expression(U[(1)],widehat(U[(1)]),widetilde(U[(1)]))), col=c("red", "green", "blue"), lty=1)
```
$U_{(1)}$ is with $m=5$, $\widehat U_{(1)}$ is with $m=20$ and $\widetilde U_{(1)}$ is with $m=100$.

It can be clearly observed that:

if $m > m'$  then $U_{(1)} \prec U_{(1)}'$

if $m < m'$  then $U_{(1)} \succ U_{(1)}'$


## Part B.

 

$F_{U_{(1)}}(x) = P(U_{(1)}<x) = 1 - P(U_{(1)}\geq x) = 1 -P(U_{1}\geq x,\ldots ,  U_{m}\geq x)$
$= 1 - P(U_{1}\geq x)*\ldots* P(U_{m}\geq x) = 1 - (1-x)^m$, $\forall x : 0\leq x\leq$. $ 1 : x\geq 1, 0 : x\leq 0$

 

This comes from independence of $U_i$, definition of $U_{(1)}$ and uniform distribution.

 

Proof of part 2:

 

$U_{(1)}\prec U_i \iff P(U_{(1)}\leq a) \geq P(U_i\leq a)$
$\iff F_{U_{(1)}}(a) \geq F_{U_{i}}(a)$
$\iff 1-(1-F_{U_{i}}(a))^{m}\geq F_{U_{i}}(a)$
$\iff 1-F_{U_{i}}(a)\geq (1-F_{U_{i}}(a))^{m}$
$\iff 1\geq (1-F_{U_{i}}(a))^{m-1}$
$\iff 1\geq 1-F_{U_{i}}(a)$
$\iff F_{U_{i}}(a)\geq 0$ $\space_{\blacksquare}$

 

Proof of part 3:

 

$U'_{(1)}\prec U_{(1)} \iff F_{U'_{(1)}}(a)\geq F_{U_{(1)}}(a)$
$\iff 1-(1-F_{U'_{i}}(a))^{m'} \geq 1-(1-F_{U_{i}}(a))^{m}$
$\iff (1-F_{U'_{i}}(a))^{m'}\leq (1-F_{U_{i}}(a))^{m}$
$\iff (1-F_{U_{i}}(a))^{m'}\leq (1-F_{U_{i}}(a))^{m}$
$\iff 1\geq (1-F_{U_{i}}(a))^{m'-m}$, $m'\geq m$
$\Longrightarrow 0\leq 1-F_{U_{i}}(a)\leq 1$
, which is always true. $_{\blacksquare}$

 

## Part C.

 

$F_{U_{(1)}}(x) = 1 - (1-F_{U_{i}}(x))^{m} = 1 - (1-x)^{m}$

 

Therefore,

 

$P(U_{(1)}\leq g(t,m)) = F_{U_{(1)}}(g(t,m)) = 1-(1-g(t,m))^{m} = t$
$\iff 1-t = (1-g(t,m))^{m}$
$\iff (1-t)^{1/m} = 1 - g(t,m)$
$\iff g(t,m)= 1 - (1-t)^{1/m}$

Therefore, the function is : $g(t,m)= 1 - (1-t)^{1/m}$.

 

## Part D.

 

$P(U_{(1)}\leq 1 - (1-t)^{1/m}) = t$
$\iff P(1-U_{(1)}\geq (1-t)^{1/m}) = t\iff$

 

$P((1-U_{(1)})^{m}\geq 1 -t) = t$
$\iff P(1-(1-U_{(1)})^{m}\leq t) = t$

 

Therefore the random variable $Y$ is given by: $Y = h_m(U_{(1)}) = 1-(1-U_{(1)})^{m}$

## Part E.


```{r}
Y <- 1-(1-vecmin_100)^100
hist(Y)
```
For $Y=1-(1-U_{(1)})^m$,

as can be seen, the random variable $Y$ that we found , holds $Y\sim Uni[0,1]$ as expected.

## Part F.

### 1.

This corresponds exactly to the function we found in part C:

$p_{value}^{adj}=1-(1-P_{value})^{\frac{1}{m}}$

where $m$ is the total number of courses (because $m=m_0$ according to the question assumptions).

So for $p_{value}=0.05$ we get:

$p_{value}^{adj}=1-(1-0.05)^{\frac{1}{m}} = 1-(0.95)^{\frac{1}{m}}$ 

As we showed and proved in previous sections, this equation is provides the adjusted pvalue, as desired. 


# Question 2

## Part 1 - Sidak's Adjusted p-values

$P_i \leq 1-(1-\alpha)^{\frac{1}{m}}=\alpha_{sid}$
$\iff P_i + (1-\alpha)^{\frac{1}{m}}\leq 1$
$\iff (1-\alpha)^{\frac{1}{m}}\leq 1-P_i$
$\iff 1-\alpha \leq (1-P_i)^m$
$\iff \alpha \geq 1-(1-P_i)^m$

Therefore,

$q_{i_{sid}} = 1-(1-P_i)^m$


## Part 2 - Simultaneous CI are FWER controlling Proc.

 

$P(V>0) \underbrace{=}_{(1)} P(\exists\space  i: \mu_{i}\notin I_{i}'(X))\underbrace{\leq}_{(2)}\alpha$

 

(1) By definition of the test.
(2) By definition of simultaneous confidence intervals.

 

Therefore, the given test is a FWER-controlling procedure at level $\alpha$.

## Part 3 - Bonferroni CI are Simultaneous CI

As seen in class, bonferroni CI for $\mu_i$ at confidence level $1-\alpha$ if it holds that $P(\mu_i \in I_i(X)) \geq 1-\frac{\alpha}{m}$ or alternatively $P(\mu_i \notin I_i(X)) < \frac{\alpha}{m}$ $(*)$

Therefore, 

$P(\exists i: \mu_i \notin I_i'(X)) = P(\bigcup\limits_{i=1}^m \{\mu_i \notin I_i'(X)\})$ 
$\underbrace{\leq}_{\mbox{bonferroni ineq.}} \sum\limits_{i=1}^m P(\mu_i\notin I_i'(X)) \underbrace{<}_{\mbox{from }(*)} \sum\limits_{i=1}^{m}\frac{\alpha}{m} = \alpha$

From definition and independence,  

$P(\forall i, \mu_i \in I_i'(X))=P(\mu_1\in I_1'(X))\cdot \ldots\cdot P(\mu_m\in I_m'(X))=(1-\frac{\alpha}{m})^{m}$

To prove that Bonferroni CI are simultaneous CI at confidence level of at least $1-\alpha$ we need to show that $(1-\frac{\alpha}{m})^{m} \geq 1 - \alpha$.

By applying the $log$ function (allowed because both sides are non negative for $m\geq 0$, $0\leq\alpha\leq1$) on both sides we get:

$(1-\frac{\alpha}{m})^{m} \geq 1 - \alpha \iff(log(1-\frac{\alpha}{m}))^{m}\geq log(1-\alpha) \iff mlog(1-\frac{\alpha}{m})\geq log(1-\alpha)$

Assuming $m\geq1$ we show that

$log(1-\frac{\alpha}{m})\geq log(1-\alpha) \iff 1-\frac{\alpha}{m}\geq 1-\alpha \iff m\geq1$.

From the assumption this is always true. $\space _{\blacksquare}$ 

## Part 4




