---
title: 'Multiple Comparisons: Homework - 2'
output:
  html_notebook: default
  github_document: 
    pandoc_args: --webtex
  pdf_document: default
---



# Question 2

## Part a.
```{r}
rec <- recovery

sum_desc=group_by(rec, blanket) %>%
  summarise(
    count = n(),
    mean = mean(minutes, na.rm = TRUE),
    sd = sd(minutes, na.rm = TRUE)
  )
sum_desc$blanket
sum_desc$mean
sum_desc$sd
sum_desc
```

```{r}
boxplot(minutes ~ blanket, data = rec,
        xlab = "Blanket", ylab = "Minutes",
        frame = FALSE, col = c("#00AFBB", "#E7B800", "#FC4E07"))
```
 
In our opinion it seems as if the recovery time when using blanket $b_2$ is different than the recovery time when using blankets $b_0, b_1$, and $b_3$.

In our opinion it seems as if the recovery time when using blanket $b_2$ is lower than the recovery time when using blanket $b_0$.

## Part b.

The assumptions in ANOVA are that

1. The data points are independent. 

2. Each population comes from a normal distribution, and all have the same standard deviation.

# TODO VERIFY FROM NOTEBOOK



```{r}

###Compute One-way ANOVA test
rec.aov <- aov(minutes ~ blanket, data = rec)
```
```{r}
#1. Homogeneity of variances

plot(rec.aov, 1)

```
```{r}
#Levene's test for equality of variances
leveneTest(minutes ~ blanket, data = rec)

```


From the results we can conclude that indeed the variances are equal and the assumption holds.

```{r}
#Check the normality assumption
# 2. Normality
plot(rec.aov, 2)
```
```{r}
#Shapiro-Wilk test on the ANOVA residuals: check normality
# Extract the residuals
aov_residuals <- residuals(object = rec.aov )
# Run Shapiro-Wilk test
shapiro.test(x = aov_residuals )
```
From the results we can conclude that indeed the normality assumption holds.



## Part c.

The Hypothesis test of ANOVA is:

$H_0 : \mu_1 = \mu_2 \cdots = \mu_k$

$H_1 : otherwise$


```{r}

# Summary of the analysis
summary(rec.aov)
```

We got F value = 7.55

P-Value = 0.000462


```{r}
###How is the p-value computed?
1-pf(7.55, 3, 37)
```

## Part d.

### 1.

Pairwise Comparisons:

$\forall \ i,j \in \{1,\cdots,k\}$

$H_{0_{(i,j)}}:\mu_i = \mu_j$

$H_{1_{(i,j)}}:\mu_i \ne \mu_j$

### 2.

For $\alpha = 0.05$:
```{r}
TukeyHSD(rec.aov, conf.level = 0.95)
```

The blankets that have a difference in terms of recovery time in expectance are:

$(b_0, b_2), (b_2, b_3)$.


For $\alpha = 0.1$:

```{r}
TukeyHSD(rec.aov, conf.level = 0.9)
```

The blankets that have a difference in terms of recovery time in expectance are:

$(b_0, b_2), (b_1, b_2), (b_2, b_3)$.
